---
title: "Analysis on Drug use"
output: html_document
---

```{r}

# setting the global setting of the chuck within this rmarkdown
knitr::opts_chunk$set(echo = TRUE, results = "hold")


# loading package required for this analysis 
if(!require(pacman)) install.packages("pacman")
p_load(
  rio, tidyverse, janitor, here, epitools, epikit, visdat, inspectdf, gtsummary, 
  tmaptools, tmap, sf, flextable, officer, pwr, tm, tidytext, wordcloud, plotly
  )



# loading data sets.
drug_value_data<-import(here("data/drug_value_chain.xlsx"))

# clean the variable names to lower cases
drug_value_data<-clean_names(drug_value_data)

# changing the characters to factors 
drug_value_data<- drug_value_data %>% 
  mutate(across(where(is.character), as.factor))

# chaning the data frame to tibble table
drug_value_data <- as_tibble(drug_value_data)


```

# Prescription and Usage Patterns
The following are variable to be analyzed under this sections: 

  
1. Receives Advice on Use?=	are_you_given_any_advice_on_their_use
2.Type of Advice Given=	if_yes_what_kind_of_advice
3.Buying Method=	what_is_the_most_frequent_way_of_buying_the_drugs
4.Action if Drug Fails=	if_you_buy_a_drug_and_it_does_not_work_what_do_you_do

```{r}
# sub-setting data of prescription and drug usage pattern 
pres_and_usage_patterns<-drug_value_data %>% 
  select(
    are_you_given_any_advice_on_their_use, 
    if_yes_what_kind_of_advice_dosage, 
    if_yes_what_kind_of_advice_duration_for_use,
    if_yes_what_kind_of_advice_route_of_administration, 
    if_yes_what_kind_of_advice_withdrawal_period,
    do_you_prescribe_drugs_to_healthy_animals_as_a_form_of_prophylaxis_315,
    if_yes_to_which_clients_poultry_farmers, 
    if_yes_to_which_clients_beef, 
    if_yes_to_which_clients_dairy_farmers, 
    if_yes_to_which_clients_pig_farmers )



pres_and_usage_patterns<-pres_and_usage_patterns %>% 
                       mutate(across(-c(1,6), ~if_else(.== 1, "Yes", "No")))


# calculating the percentage by row of the values.
pres_and_usage_patterns<- pres_and_usage_patterns %>% 
       pivot_longer(
         cols = 1:10, 
         names_to = "Category"
       ) %>% 
  drop_na(value) %>% 
  group_by(Category, value) %>% 
  summarise(freq.=n()) %>% 
  pivot_wider(
    values_from = freq., 
    names_from = value
  ) %>% 
  mutate(
    total=round(
      No+Yes
    ), 
    Yes=round(
      Yes/total*100, 2
    ), 
    No=round(
      No/total*100, 2
    )
  ) %>% 
  select(
    Category, Yes, No
  ) %>% 
  pivot_longer(
    cols = -Category
  )


# plotting chart of drug use and patterns 

pres_and_usage_patterns$Category <- gsub("_", " ", 
                                pres_and_usage_patterns$Category)



pres_and_usage_patterns$Category <- gsub("if yes to which clients",
                                         "prescribe drugs to", 
                                pres_and_usage_patterns$Category)


pres_and_usage_patterns$Category <- gsub("if yes what",
                                         " ", 
                                pres_and_usage_patterns$Category)


# Plotting the bar chart. 
pres_usage_plot<- pres_and_usage_patterns %>% 
    dplyr::rename("Response"=name) %>% 
  ggplot( aes(x=Category, y= value, fill = Response))+
  coord_flip()+
  geom_col()+
  theme_minimal()+
  theme(axis.text.x = element_text(
    angle = 0, size = 12, colour = "black" ), 
    axis.text.y = element_text(
      size = 13, colour = "black" 
    ))+
  geom_text(  aes(
           x=Category, y=value, 
           label = value),hjust =0 , vjust = 0.9  
            )+
  labs(
    y="Percentage", 
  title= "Prescription and Usage Patterns" )


# saving the image 
ggsave(filename = "pres_usage_plot.png", 
       plot = pres_usage_plot,
       path = "output", bg="white", dpi = 300)




```

## Awareness of policies and institutions
are_you_aware_of_the_institutions_that_govern_drug_use
are_you_aware_of_any_policies_governing_drug_use_storage_handling
where_did_you_get_this_information_from
do_you_think_that_the_existing_policies_are_effective
if_no_what_should_be_done


```{r}
# Awareness of the policies 
awareness_of_policies <- drug_value_data %>% 
              select(331, 333, 336:338, 340,341, 343) 

# re-coding all the value indicate 1 to Yes, and 0 to No exclude column with yes, and no values
awareness_of_policies<-awareness_of_policies %>% 
  mutate(across(-c(8,7, 1,2), ~if_else(.== 1, "Yes", "No")))

# calculating row percentage 
awareness_of_policies<- awareness_of_policies %>% 
  pivot_longer(
    cols = 1:8, 
    names_to = "Category"
  ) %>% 
  drop_na(value) %>% 
  group_by(Category, value) %>% 
  summarise(freq.=n()) %>% 
  pivot_wider(
    names_from = value, 
    values_from = freq.) %>% 
  mutate( total=round( No+Yes),
          Yes=round(Yes/total*100, 2), 
          No=round(No/total*100, 2)) %>% 
  select(Category, Yes, No) %>% 
  pivot_longer(
    cols = -Category
  ) 
  
# recoding values within the category variable      
awareness_of_policies<- awareness_of_policies %>% 
  mutate(Category =recode(
    Category, 
      "are_you_aware_of_the_institutions_that_govern_drug_use_331"=
      "Aware of the institution that govern drug use", 
      "do_you_think_that_the_existing_policies_are_effective_343"=
      "The existing policies are effective", 
      "where_did_you_get_this_information_from_background_training_338" = 
      "I get the information from background training",
      "where_did_you_get_this_information_from_education_training_340"= 
      "I get the information from Education training", 
      "where_did_you_get_this_information_from_internet_press_336"= 
      "I get the information from internet press", 
      "where_did_you_get_this_information_from_workshops_trainings_337"=
      "I get the information from Workshop", 
      "are_there_existing_channels_for_information_sharing_between_you_and_the_institutions_341"= 
      "Existing channels for information sharing between a farmer and institutions ", 
   "are_you_aware_of_any_policies_governing_drug_use_storage_handling_333"=
   "Awareness of policies governing drug use, storage and handling"
    ))

awareness_of_policies$Category <- str_wrap(awareness_of_policies$Category, 
                                           width = 40)


# plotting a chart

awareness_of_policies_plot<- awareness_of_policies %>% 
    dplyr::rename("Response"=name) %>% 
  ggplot( aes(x=Category, y= value, fill = Response))+
  coord_flip()+
  geom_col()+
  theme_minimal()+
  theme(axis.text.x = element_text(
    angle = 0, size = 8, colour = "black" ), 
    axis.text.y = element_text(
      size = 13, colour = "black" 
    ))+
  geom_text(
    aes(label = value),  # Show actual percentage
    position = position_stack(vjust = 0.5),  # Center text inside bars
    color = "black",  # Ensure visibility inside bars
    size = 4
  ) +
  scale_fill_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +  # Set fill colors
  labs(y = "Percentage", title = "Awareness of Policies")


# saving the image 
ggsave(filename = "awareness_of_policies_plot.png", 
       plot = awareness_of_policies_plot,
       path = "output", bg="white", dpi = 300, width = 10, height = 6)




```

## Ways of making policies effective


```{r}


# creating a subset of data set
ways_of_policies_effective<-drug_value_data %>% 
                           select(345:347)

ways_of_policies_effective<-ways_of_policies_effective %>% 
  mutate(across(everything(), ~if_else(. ==1, "Yes", "No")))


ways_of_policies_effective<- ways_of_policies_effective %>% 
  pivot_longer(
    cols = 1:3, 
    names_to = "Category"
  ) %>% 
  drop_na(value) %>% 
  group_by(Category, value) %>% 
  summarise(freq.=n()) %>% 
  pivot_wider(
    names_from = value, 
    values_from = freq.) %>% 
  mutate(total=round(No+Yes), 
         Yes=round(Yes/total*100, 2), 
         No=round(No/total*100, 2)) %>% 
  select(Category, Yes, No) %>% 
  pivot_longer(
    cols = -Category
  )
  
ways_of_policies_effective$Category <- gsub("if_no_what_do_you_think_can_be_done_to_make_them_more_effective_", 
     "",ways_of_policies_effective$Category)

ways_of_policies_effective$Category <- gsub("_", 
     " ",ways_of_policies_effective$Category)

## plotting the chart

ways_of_policies_plot<- ways_of_policies_effective %>% 
    dplyr::rename("Response"=name) %>% 
  ggplot( aes(x=Category, y= value, fill = Response))+
  coord_flip()+
  geom_col()+
  theme_minimal()+
  theme(axis.text.x = element_text(
    angle = 0, size = 8, colour = "black" ), 
    axis.text.y = element_text(
      size = 13, colour = "black" 
    ))+
  geom_text(
    aes(label = value),  # Show actual percentage
    position = position_stack(vjust = 0.5),  # Center text inside bars
    color = "black",  # Ensure visibility inside bars
    size = 4) +
  scale_fill_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +  # Set fill colors
  labs(y = "Percentage", title = "Way of Improve Awareness of Policies")



# saving the image 
ggsave(filename = "ways_of_policies_plot.png", 
       plot = ways_of_policies_plot,
       path = "output", bg="white", dpi = 300)

```



Dumbbell Chart
Shows the gap between "Yes" and "No

```{r}

# 
# ggplot(ways_of_policies_effective, aes(x = Category, y = value, color = name)) +
#   geom_point(size = 5) +
#   geom_line(aes(group = Category), size = 1) +
#   coord_flip() +
#   theme_minimal() +
#   scale_color_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +
#   labs(y = "Percentage", title = "Way of Improve Awareness of Policies")


```


 Lollipop Chart
Emphasizes values with points and lines

```{r}
# 
# 
# ggplot(ways_of_policies_effective, aes(x = Category, y = value, color = name)) +
#   geom_segment(aes(xend = Category, y = 0, yend = value), size = 1) +
#   geom_point(size = 5) +
#   coord_flip() +
#   theme_minimal() +
#   scale_color_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +
#   labs(y = "Percentage", title = "Way of Improve Awareness of Policies")

```


Transform "No" values to negative for a clear contrast

```{r}
# 
# ways_of_policies_effective2 <- ways_of_policies_effective %>%
#   mutate(value = ifelse(name == "No", -value, value))
# 
# ggplot(ways_of_policies_effective2, aes(x = Category, y = value, fill = name)) +
#   geom_col() +
#   coord_flip() +
#   theme_minimal() +
#   scale_fill_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +
#   labs(y = "Percentage", title = "Way of Improve Awareness of Policies") +
#   geom_text(aes(label = abs(value)), position = position_stack(vjust = 0.5))


```


Dodge Bar Chart with Percent Labels
Like the grouped bar chart but with labels:

```{r}

# 
# ggplot(ways_of_policies_effective, aes(x = Category, y = value, fill = name)) +
#   geom_col(position = position_dodge(width = 0.8)) +
#   geom_text(aes(label = value), position = position_dodge(width = 0.8), vjust = -0.5) +
#   coord_flip() +
#   theme_minimal() +
#   scale_fill_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +
#   labs(y = "Percentage", title = "Way of Improve Awareness of Policies")

```


## Stakeholder Engagement

dis_attend_any_stakeholder_meetings_the_last_6_months
if_yes_who_were_the_organizers
which_other_stakeholders_attend_such_meetings
if_you_dont_attend_why

```{r}

stakeholder_engagement <- drug_value_data %>% 
        select(349, 353)

stakeholder_engagement<- stakeholder_engagement %>% 
  pivot_longer(
    cols = 1:2, 
    names_to = "Category"
  ) %>% 
  drop_na(value) %>% 
  group_by(Category, value) %>% 
  summarise(freq.=n()) %>% 
  pivot_wider(
    names_from = value, 
    values_from = freq.) %>% 
  mutate(total=round(No+Yes), 
         Yes=round(Yes/total*100,2), 
         No=round(No/total*100, 2)) %>% 
  select(Category, Yes, No) %>% 
  pivot_longer(
    cols = -Category
  ) %>% 
  mutate(Category =recode(
    Category, 
    "do_you_attend_any_stakeholder_meetings_349"="Attended any Stakehodler meeting", 
  "do_you_think_there_is_sufficient_interaction_between_you_and_the_regulating_bodies_353"="Interaction between you and Regulating bodies"
  ))



# # plotting a pie chart 
# 
# library(ggplot2)
# library(dplyr)
# library(scales)  # For percentage formatting

# Create the pie chart
stakeholder_plot<- ggplot(stakeholder_engagement, aes(x = "", y = value, fill = name)) +
  geom_bar(stat = "identity", width = 1, color = "white") +  # White borders for clarity
  coord_polar(theta = "y") +  # Convert to pie chart
  facet_wrap(~ Category) +  # Separate pies for each category
  theme_void() +  # Remove unnecessary chart elements
  geom_text(aes(label = paste0(round(value, 1), "%")), 
            position = position_stack(vjust = 0.5), size = 5, color = "black") +  # Add labels inside the pie
  scale_fill_manual(values = c("Yes" = "skyblue", "No" = "tomato")) +
  labs(title = "Stakeholder Engagement ")


# saving the image 
ggsave(filename = "stakeholder_plot.png", 
       plot = stakeholder_plot,
       path = "output", bg="white", dpi = 300)




```



