---
title: "Analysis of the AMU data"
output:
  html_document:
    df_print: paged

---

```{r , results='hold'}
# setting the global setting of the chuck within this rmarkdown
knitr::opts_chunk$set(echo = TRUE, results = "hold")


# loading package required for this analysis 
if(!require(pacman)) install.packages("pacman")
p_load(
  rio, tidyverse, janitor, here, epitools, epikit, visdat, inspectdf, gtsummary, 
  tmaptools, tmap, sf, flextable, officer
)


# packages for qauilative analysis in R

p_load(
  tm, tidytext, wordcloud, plotly
)


# loading data sets.
amu_data<-import(here("data/AMU.xlsx"))

# clean the variable names to lower cases
amu_data<-clean_names(amu_data)

# changing the characters to factors 
amu_data<- amu_data %>% 
  mutate(across(where(is.character), as.factor))

options(flextable.print = "full")



```

# descripitive stasticts for demographic variables 

```{r}
# sub setting demographic variables.
demo_data<-amu_data %>% 
  select(age_of_the_participant,
        gender_of_the_participant, 
        marital_status_of_the_participant, 
        level_of_education_of_the_participant,
        role_of_the_participant_in_livestock_production,
        # village, # remove village because it not clean 
        # ta, # remove ta variable because the values are not clean
        district) %>% 
  mutate(age_group=age_categories(
        age_of_the_participant, 
        breakers = c(0,35, 45, 55)
  )) %>% 
  select(-age_of_the_participant) #

# re-arranging the variables
demo_data<- demo_data %>% 
  select(age_group, everything())

# summary of the demographic variables
summary(demo_data)

# creating freq. table of the demographic variables
demo_table<-demo_data %>% 
  tbl_summary()


# converting the gt object to flex table object

demo_table <-  as_flex_table(demo_table) 
               
# saving the demographic table

save_as_docx(demo_table, path = here("output/demographics_table.docx"))


```

# Descriptive stat. farm characteristics

```{r}
# creating a sub set of livestock raise in the farm
livestock_raise<-amu_data %>% 
  dplyr::select(c(24: 28)) %>% 
  rename(
    "Poultry"= type_of_livestock_raised_poultry_layers_broilers_indigenous, 
    "Dairy_cattle"= type_of_livestock_raised_dairy_cattle, 
    "Beef_Cattle"=type_of_livestock_raised_beef_cattle, 
    "Pigs"=type_of_livestock_raised_pigs, 
    "Goats"=type_of_livestock_raised_goats
  )

# creating a freq. table of livestock keep at the farm

tabel_livestoc_raise<-livestock_raise %>% 
  pivot_longer(cols = 1:5, 
               names_to = "Category") %>% # changing the data structure
  filter(value == 1) %>% # filtering out those that responsed or tick 
  group_by(Category) %>% # grouping the data by category
  summarise(n = n()) %>% # summaring the data by counts of category
  mutate(Percentage = round(n / sum(n) * 100, 3)) # calculating percentage

# sub setting number of livestock keep on the farm
number_livestoct <- amu_data %>% 
        dplyr::select( c(31:35))



# creating a table on the number of liverstock keep at the farm
table_number_livestoct<-number_livestoct %>% 
  pivot_longer(
    cols = 1:5, 
    names_to = "Category"
  ) %>% 
  filter(value>=1) %>% 
  group_by(Category) %>% 
  summarise(n=sum(value)) %>% 
  mutate(Percentage=round(
    n/sum(n)*100, 3
  ))


# subsetting varaible number 37 to 39
table1_livestock<-amu_data %>% 
  select(c(37:39))

# creating a table on number of years in expereience of keeoing livestock
table1_livestock<- table1_livestock %>% 
                       tbl_summary()

# converting the livestock table to flextable object

table1_livestock<-as_flex_table(table1_livestock)

# mergeing the two first tables 
merge_table<-union_all(
 tabel_livestoc_raise, 
 table_number_livestoct
)

# converting the table to flextable object

table_livestock_2<-as_flextable(merge_table)


# saving the tables 

## saving table 1 
save_as_docx(table_livestock_2, path = here("output/table.docx"))

save_as_docx(table1_livestock, path = here("output/table_2.docx"))


```


# Antimicrobial use practices


```{r}
options(flextable.shadow = FALSE)

# subsetting antimicrobial use data set
antimicrobial <- amu_data %>% 
    select(41: 81) %>% 
    select(-matches(("other_specify|specify"))) %>%  # deselect all the variables contain others and specify.
    select(-c(where_do_you_purchase_antimicrobials, 
           how_do_you_administer_antimicrobials_to_your_livestock_select_all_that_apply, 
          if_yes_can_you_show_us_the_bottles_packages_for_verification_url, 
          if_yes_can_you_show_us_the_bottles_packages_for_verification, 
      can_you_provide_the_names_of_antimicrobials_currently_used_on_your_farm,
      how_do_you_decide_which_antimicrobial_to_use,
      what_are_the_main_reasons_for_using_antimicrobials, 
      if_no_explain_why_72, 
      how_do_you_store_antimicrobials_on_your_farm, 
      how_do_you_dispose_of_unused_or_expired_antimicrobials)
      )  



# Convert all variables from 0="No", 1= "Yes"

antimicrobial<- antimicrobial %>% 
   mutate(across(-c(1,9,16), ~ ifelse(. == 1, "Yes", "No")))

antimicrobial<-as_tibble(antimicrobial)

# change the data structure

table3<-antimicrobial %>% 
  pivot_longer(
    cols = 1:21
  ) %>% 
  filter(value=="Yes"|value=="No") %>% 
  group_by(name, value) %>%
  count(value) %>%
  pivot_wider(
    names_from = value,
  values_from = n
  )

# replacing NA with 0
table3[is.na(table3)]<-0



# Creating a table on the Antimicrobial use practices
table3<-table3 %>% 
  mutate(total=round(
    Yes+No
  ), 
  Yes=round(
    Yes/total*100, 3
  ), 
  No=round(
    No/total*100, 3
  )) %>% 
  rename(
    "Variable"=name
  ) %>% 
  select(
    Variable, 
    Yes, 
    No
  )



# replacing _ with space in the column variable
 table3$Variable <- gsub("_", " ", table3$Variable) 
  

#save the table as an excel 

export(table3, "output/table3.xlsx")




```

# Disease management practices

```{r}

disease_management<-amu_data %>% 
                       select(
                        84:98  )

disease_management<- disease_management %>% 
  mutate(across(where(is.factor), as.character))


```

## Plotting disease in poulty 

```{r}

poultry_disease<-  disease_management %>% 
              unnest_tokens( poultry_disease,
                diseases_of_poultry_currently_on_the_farm) %>% 
                select(poultry_disease) %>% 
        cross_join(stop_words)


poultry_disease<-poultry_disease %>% 
                drop_na(poultry_disease)


count_words <- poultry_disease %>% 
      count(poultry_disease, sort = T)





# plotting a word cloud for poultry 

output_path <- here("output", "Poultry_Disease.png")

# **Save Word Cloud as PNG**
png(output_path, width = 800, height = 600)
  


wordcloud(words = count_words$poultry_disease, 
          freq = count_words$n, 
          min.freq = 1,          # Minimum frequency of words to be included
          max.words = 100,       # Maximum words to be displayed
          random.order = FALSE,  # Arrange words based on frequency
          rot.per = 0.35,    # Proportion of words rotated (35% at 90 degrees)
          colors = brewer.pal(8, "Dark2")) # Color palette

dev.off() 

```

## Disease in Dailry cattle

```{r}

Dairy_disease<-  disease_management %>% 
              unnest_tokens(Dairy_disease,
                diseases_of_dairy_cattle_currently_on_the_farm) %>% 
                select(Dairy_disease) %>% 
        cross_join(stop_words)


Dairy_disease<-Dairy_disease %>% 
                drop_na(Dairy_disease)


count_words <- Dairy_disease %>% 
      count(Dairy_disease, sort = T)





# plotting a word cloud for poultry 

output_path <- here("output", "Dairy_Disease.png")

# **Save Word Cloud as PNG**
png(output_path, width = 800, height = 600)
  


wordcloud(words = count_words$Dairy_disease, 
          freq = count_words$n, 
          min.freq = 1,          # Minimum frequency of words to be included
          max.words = 100,       # Maximum words to be displayed
          random.order = FALSE,  # Arrange words based on frequency
          rot.per = 0.35,    # Proportion of words rotated (35% at 90 degrees)
          colors = brewer.pal(8, "Dark2")) # Color palette

dev.off() 



```

### Disease of Beef cattle 

```{r}

Beef_disease<-disease_management %>% 
              unnest_tokens(Beef_disease,
                diseases_of_beef_cattle_currently_on_the_farm) %>% 
                select(Beef_disease) %>% 
        anti_join(stop_words)


Beef_disease<-Beef_disease %>% 
                drop_na(Beef_disease)


count_words <- Beef_disease %>% 
      count(Beef_disease, sort = T)


# plotting a word cloud for poultry 

output_path <- here("output", "Beef_Disease.png")

# **Save Word Cloud as PNG**
png(output_path, width = 600, height = 400)
  


wordcloud(words = count_words$Beef_disease, 
          freq = count_words$n, 
          min.freq = 1,          # Minimum frequency of words to be included
          max.words = 100,       # Maximum words to be displayed
          random.order = FALSE,  # Arrange words based on frequency
          rot.per = 0.35,    # Proportion of words rotated (35% at 90 degrees)
          colors = brewer.pal(8, "Dark2")) # Color palette

dev.off() 



```

## Disease in Pigs 

```{r}

Pig_disease<-disease_management %>% 
              unnest_tokens(Pig_disease,
                diseases_of_pigs_currently_on_the_farm) %>% 
                select(Pig_disease)%>% 
        cross_join(stop_words)


Pig_disease<-Pig_disease %>% 
                drop_na(Pig_disease)


count_words <- Pig_disease %>% 
      count(Pig_disease, sort = T)


# plotting a word cloud for poultry 

output_path <- here("output", "Pig_disease.png")

# **Save Word Cloud as PNG**
png(output_path, width = 600, height = 400)
  


wordcloud(words = count_words$Pig_disease, 
          freq = count_words$n, 
          min.freq = 1,          # Minimum frequency of words to be included
          max.words = 100,       # Maximum words to be displayed
          random.order = FALSE,  # Arrange words based on frequency
          rot.per = 0.35,    # Proportion of words rotated (35% at 90 degrees)
          colors = brewer.pal(8, "Dark2")) # Color palette

dev.off() 


```

## Goat Disease 

```{r}

Goat_disease<-disease_management %>% 
              unnest_tokens(Goat_disease,
                diseases_of_goats_currently_on_the_farm) %>% 
                select(Goat_disease)%>% 
        cross_join(stop_words)


Goat_disease<-Goat_disease %>% 
                drop_na(Goat_disease)


count_words <- Goat_disease %>% 
      count(Goat_disease, sort = T)


# plotting a word cloud for poultry 

output_path <- here("output", "Goat_disease.png")

# **Save Word Cloud as PNG**
png(output_path, width = 8, height = 5, units = "cm", res = 300)

wordcloud(words = count_words$Goat_disease, 
          freq = count_words$n, 
          min.freq = 1,          # Minimum frequency of words to be included
          max.words = 100,       # Maximum words to be displayed
          random.order = FALSE,  # Arrange words based on frequency
          rot.per = 0.35,    # Proportion of words rotated (35% at 90 degrees)
          colors = brewer.pal(8, "Dark2")) # Color palette

dev.off() 



```

## Common methods of treating disease 


```{r}
# sub setting common treatment within the disease management data set
animal_treatment <- disease_management %>% 
                  select(
                    starts_with("how_do_you_usually") & # select var. with how
                    - ends_with(c ("other_specify","all_that_apply") #         removing variables end with others and all that apply
                  ))


# Convert all variables from 0="No", 1= "Yes"

animal_treatment<- animal_treatment %>% 
   mutate(across(everything(), ~ ifelse(. == 1, "Yes", "No")))


animal_treatment<- animal_treatment %>% 
  pivot_longer(
    cols = 1:4, 
    names_to = "Category"
  ) %>% 
  drop_na(value) %>% 
  group_by(
    Category, 
    value
  ) %>% 
  summarise(freq.=n())

# replace the phase "how_do..._" with space 
animal_treatment$Category<- gsub("how_do_you_usually_treat_these_diseases_select_all_that_apply_",
     " " ,
     animal_treatment$Category)

# replacing _ with space
animal_treatment$Category<- gsub("_", " " , animal_treatment$Category)

# restructuring the data set for plot a stock bar chart 
animal_treatment<- animal_treatment %>% 
  pivot_wider(
    names_from = value, 
    values_from = freq.
  ) %>% 
  mutate(
    total=round(
      No+Yes
    ), 
    Yes=round(
      Yes/total*100, 2
    ), 
    No=round(
      No/total*100,2
    )
  ) %>% 
  select(
    Category, 
    Yes, 
    No
        ) %>% 
  pivot_longer(
    cols = -1, 
    
  )

# creating a chart 

treatment_plot<-animal_treatment %>% 
  ggplot( aes(x=Category, y= value, fill = name))+
  coord_flip()+
  geom_col()+
  theme_minimal()+
  theme(axis.text.x = element_text(
    angle = 0, size = 8, colour = "black"
  ))+
  geom_text(  aes(
           x=Category, y=value, 
           label = value))+
  labs(
    y="Percentage", 
    title = "Methods how Animal Disease are Treated "
  )


# saving the image 
ggsave(filename = "treatment_plot.png", plot = treatment_plot,
       path = "output", bg="white", dpi = 300)

```

## Vaccinate uage 

```{r}
vaccine <-disease_management %>% 
          select(
            are_vaccines_used_on_your_farm
          ) 


vaccine_pie<- vaccine %>% 
   group_by(are_vaccines_used_on_your_farm) %>% 
   count(vaccine$are_vaccines_used_on_your_farm)%>%
   drop_na(are_vaccines_used_on_your_farm) %>% 
  summarise(count=n/n()) %>% 
  mutate(precent=  round(count/sum(count)*100, 3)) %>% 
  select(-count) %>% 
  rename(
    "vacination_used"=are_vaccines_used_on_your_farm
  )


# restructing the data set for pie chart
vaccine_pie2<- vaccine_pie %>% 
mutate(csum = rev(cumsum(rev(precent))),
        pos = precent/2 + lead(csum, 1),
        pos = if_else(is.na(pos), precent/2, pos))

# plotting the pie chart
pie_chart<- ggplot(vaccine_pie2, aes(x = "", y = precent,
        fill = vacination_used)) + 
      geom_bar(stat="identity",color=1) + 
      coord_polar(theta="y",start=0) + 
      scale_fill_brewer() + 
      ggtitle("Is Vaccine Use at Farm") + 
      guides(fill=guide_legend(title="Vaccine used")) + 
       geom_text(aes(label = paste0(round(precent, 1), "%"), 
       y = pos), color = "black", size = 5)+
      theme_void() + theme(legend.position="right")


# saving the pie chart 
ggsave(plot = pie_chart, path = here("output"), 
       filename ="pie_chart.png")

```


## Knowledge and awareness of antimicrobial resistance 
### awareness of amr

```{r}
knowledge_awareness<-amu_data%>% 
              select(
                112:137
              ) 

awareness_amr <- knowledge_awareness %>% 
         select( have_you_heard_of_amr)


amr_pie<- awareness_amr %>% 
   group_by(have_you_heard_of_amr) %>% 
   count(have_you_heard_of_amr)%>%
   drop_na(have_you_heard_of_amr) %>% 
  summarise(count=n/n()) %>% 
  mutate(precent=  round(count/sum(count)*100, 3)) %>% 
  select(-count) %>% 
  rename(
    "Awareness of AMR"=have_you_heard_of_amr
  )




# restructing the data set for pie chart
amr_pie<- amr_pie %>% 
mutate(csum = rev(cumsum(rev(precent))),
        pos = precent/2 + lead(csum, 1),
        pos = if_else(is.na(pos), precent/2, pos))


# plotting the pie chart
AMR_chart<- ggplot(amr_pie, aes(x = "", y = precent,
        fill = `Awareness of AMR`)) + 
      geom_bar(stat="identity",color=1) + 
      coord_polar(theta="y",start=0) + 
      scale_fill_brewer() + 
      ggtitle("Awareness of AMR") + 
      guides(fill=guide_legend(title="AMR")) + 
       geom_text(aes(label = paste0(round(precent, 1), "%"), 
       y = pos), color = "black", size = 5)+
      theme_void() + theme(legend.position="right")


# saving the pie chart 
ggsave(plot =AMR_chart, path = here("output"), 
       filename ="AMR_chart.png")



```

### AMR policy, and training. 

```{r}
# sub setting awareness of policy and training status on AMR
amr_policy_training <- knowledge_awareness %>% 
      select(
        c(3, 5, 7, 8, 9, 12, 14),17:23 
      )

amr_policy_training<- amr_policy_training %>% 
  mutate(across(-c(1,2,6,7), ~ ifelse(. == 1, "Yes", "No")))


amr_policy_training<- amr_policy_training %>% 
  pivot_longer(
    cols = 1:14, 
    names_to = "Category"
  ) %>% 
  drop_na(value) %>% 
  group_by(
    Category, 
    value
  ) %>% 
  summarise(freq.=n())


# replace the phase "if_yes_..._" with space 
amr_policy_training$Category<- gsub("if_yes_who_provides_this_",
     " " ,
     amr_policy_training$Category)



# replace _ with space 
amr_policy_training$Category<- gsub("_",
     " " ,
     amr_policy_training$Category)


# restructuring the data set for plot a stock bar chart 
amr_policy_training<- amr_policy_training %>% 
  pivot_wider(
    names_from = value, 
    values_from = freq.
  ) %>% 
  mutate(
    total=round(
      No+Yes
    ), 
    Yes=round(
      Yes/total*100, 2
    ), 
    No=round(
      No/total*100,2
    )
  ) %>% 
  select(
    Category, 
    Yes, 
    No
        ) %>% 
  pivot_longer(
    cols = -1, 
    
  )


# recoding the values of the category 
amr_policy_training<-amr_policy_training%>% 
  mutate(Category= recode(
    Category, 
    "are you aware of any guidelines or regulations on antimicrobial use in livestock"= "Awareness of Guideline/Regulation on Antimicrobial", 
  "do you receive any training or advice on responsible antimicrobial use"= "Trained or Adviced on Antimicrobial", 
  " training friends through some discussions"="Trained on AMR by Friends",
  " training government through veterinarians"= "Trained by Government Veterianrians", 
  " training ngo"= "Trained on AMR by NGO"
  ))


# plotting stock bar chart 

amr_policy_plot<-amr_policy_training%>% 
  dplyr::rename(
    "Response"=name
  ) %>% 
  ggplot( aes(x=Category, y= value, fill = Response))+
  coord_flip()+
  geom_col()+
  theme_minimal()+
  theme(axis.text.x = element_text(
    angle = 0, size = 12, colour = "black" ), 
    axis.text.y = element_text(
      size = 13, colour = "black" 
    ))+
  geom_text(  aes(
           x=Category, y=value, 
           label = value))+
  labs(
    y="Percentage", 
    title = "Awareness of AMR Policy and AMR training "
  )


# saving the image 
ggsave(filename = "amr_policy_plot.png", plot = amr_policy_plot,
       path = "output", bg="white", dpi = 300)



```


